---
title: "lm_accident_prediction"
author: "Ali_Mirzaei"
date: "2022-11-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r library}
library('caret')
```
#data explantory
```{r data}
glimpse(df)
```


#plot Accidents for 365 dayes
```{r plot accidents}
plot(df$acc_total, main = "Time series",type = "l")
```
#lm model0 
```{r lm}
modl0 <- lm(acc_total ~ 
              TOTAL_VEHICLES, data=df)
summary(modl0 )
```
#residuals
```{r residuals}
plot(fitted(modl0),residuals(modl0), xlab="fitted values",ylab="residuals")
```
#confidence interval for coefficients
```{r confidence interval}
confint(modl0, 'TOTAL_VEHICLES', level=0.95)
```
#pairs plot 
```{r pairs plot}
corrplot(
  cor(df %>% select(-JDATE,-JYEAR, -TAG,
                    -JMONTH, -JDAY, -JDATE,
                    -IS_HOLIDAY, -DESCRIPTION,
                   )),
  method = 'square',
  type = 'upper',
  tl.col = 'black',
  tl.cex = 0.75, tl.srt = 90,
  col = colorRampPalette(c('purple', 'dark green'))(200)
)
```
#model matrix
```{r model matrix}
model.matrix(modl0)
```
#model plot
```{r model plot}
par(mfrow=c(2,2)) # split the graphics device into 4 panels
plot(modl0) # (uses plot.lm as modl0 is class ‘lm’)
```
#AIC 
```{R AIC }
AIC(modl0)
```
#model selection 
##lm modl1
```{r lm1}
modl1 <- lm(acc_total ~ 
              SPEED_VIOLATIONS, data=df)
summary(modl1 )
```
##AIC FOR TWO MODEL
```{r model selection}
AIC(modl1,modl0)
```
# Specifying an 80-20 train-test split
```{r test data}
train_idx = createDataPartition(df$SPEED_VIOLATIONS, p = .8, list = F)
train = df[train_idx, ]
test = df[-train_idx, ]
print(test)
```
# Declaring the trainControl function
```{r trainControl function}
train_ctrl = trainControl(
  method  = "cv", #Specifying Cross validation
  number  = 2, # Specifying 5-fold
  verboseIter = TRUE, # So that each iteration you get an update of the progress
  classProbs = TRUE # So that you can obtain the probabilities for each example
)
rf_model = train(
  acc_total ~SPEED_VIOLATIONS, # Specifying the response variable and the feature variables
  method = "rf", # Specifying the model to use
  data = train, 
  trControl = train_ctrl
)
```

## Get the predictions of your model in the test set 
```{r pred}
predictions = predict(rf_model, newdata = test)
print(predictions)
```
## factor variables
Notice how R reports an intercept parameter and parameters for the two treatment
levels, but, in order to obtain an identifiable model, it has not included a parameter
for the control level of the group factor.
TAG is a factor variable in my data.
```{r factor}
df$TAG <- as.factor(df$TAG)
fit <- lm(acc_total ~ TAG , data=df)
summary(fit)
par(mfrow=c(3,3))
plot(fit) # Then R will show you four diagnostic plots one by one:1. Residuals vs Fitted,2. Normal Q-Q,3. Scale-Location,4. Residuals vs Leverage
```
## Measures of Influence
```{r Influence}
library(olsrr)
par(mfrow=c(1,6))
ols_plot_cooksd_bar(modl1)
ols_plot_dfbetas(modl1)
ols_plot_resid_stud(modl1)
ols_plot_resid_lev(modl1)
ols_plot_resid_stud_fit(modl1)
ols_plot_resid_pot(modl1)
```
#confint
```{r confidence interval}
confint(modl1, 'SPEED_VIOLATIONS', level=0.95)
```


